---
title: "LESO wrangling"
date: "`r Sys.Date()`"
---

::: callout-warning
March 31, 2023 update. With each update, check the "Austin MSA agencies" in the analysis notebook.
:::

This notebook processes [data](https://www.dla.mil/DispositionServices/Offers/Reutilization/LawEnforcement/PublicInformation/) by the Defense Logistics Agency about military surplus transfers to local law enforcement through the Law Enforcement Support Office (LESO) or [LESO Program](https://www.dla.mil/DispositionServices/Offers/Reutilization/LawEnforcement/).

The data is updated quarterly. As of this date, the file name is linked from the headline "ALASKA - WYOMING AND US TERRITORIES".

The spreadsheet has a a sheet for each state. This notebook puts all the sheets together into a single dataset for easier analysis.

There is some cleaning done here based on June 2022 research. See the README.

## Setup

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(readxl)
library(janitor)
```

## Import

```{r import}
# set path to data
## change name to update to new file
path <- "data-raw/AllStatesAndTerritoriesQTR4FY23.xlsx"

# import and combine sheets
leso <- path %>%
  excel_sheets() %>%
  map_df(~ read_excel(path = path, sheet = .x), .id = "sheet") %>% 
  clean_names()

# peek at names
leso %>% names()
```

```{r peek}
# peek at df
leso %>% head()

```

```{r glimpse}
# glimpse at df
leso %>% glimpse()
```

```{r}
leso %>% summary()
```

## Clean up RECON SCOUT XT,SPEC

At one point, I found a mistake in the data where a recon robit was not considered controlled. I questioned DLA about it and they said it was a mistake. I fixed it in that version of the data, but it isn't in the most recent version. It is worth checking.

```{r}
leso |> 
  filter(str_detect(item_name, "RECON SCOUT")) |> 
  count(item_name, demil_code)
```

If anything in that list shows up as `demil_code` "A" or "Q6" then it should be changed to "D".

You can use the code below to change them. It's commented out for now. Update the export tibble if needed.

```{r}
# leso_fixed <- leso |> 
#   mutate(
#     demil_code = case_when(
#       item_name == "RECON SCOUT XT,SPEC" ~ "D",
#       TRUE ~ demil_code
#     )
#   )
```


## Some basic checks

This just makes sure we didn't miss a bunch of states.

```{r}
leso |> 
  count(state)
```


## Write files

Write files for export for future notebooks or other reasons.

```{r export}

# write data to csv
leso %>% 
  write_csv("data-processed/leso.csv")

leso %>% 
  write_rds("data-processed/leso.rds")
```


