---
title: "LESO data wrangling"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

This notebook processes [data](https://www.dla.mil/DispositionServices/Offers/Reutilization/LawEnforcement/PublicInformation/) by the Defense Logistics Agency about military surplus transfers to local law enforcement through the Law Enforcement Support Office (LESO) or [LESO Program](https://www.dla.mil/DispositionServices/Offers/Reutilization/LawEnforcement/).

The data is updated quarterly. As of this date, the file name is linked from the headline "ALASKA - WYOMING AND US TERRITORIES".

The spreadsheet as a a sheet for each state. This notebook puts all the sheets together into a single dataset for easier analysis.

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(readxl)
library(janitor)
```

## Import

```{r import}
# set path to data
## change name to update to new file
path <- "data-raw/DISP_AllStatesAndTerritories_09302020.xlsx"

# import and combine sheets
leso <- path %>%
  excel_sheets() %>%
  map_df(~ read_excel(path = path, sheet = .x), .id = "sheet") %>% 
  clean_names()

# peek at names
leso %>% names()
# peek at df
leso %>% head()
```

## Texas

Filter for Texas agencies

```{r texas}
tx <- leso %>% 
  filter(state == "TX")

```

## Write files

Write files for export for future notebooks or other reasons.

```{r export}

# write data to csv
leso %>% 
  write_csv("data-processed/leso.csv")

leso %>% 
  write_rds("data-processed/leso.rds")

# klaxon on complete
beepr::beep(4)
```

