---
title: "LESO Texas and Austin"
date: "`r Sys.Date()`"
---

## Setup

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
```

## Import

```{r import}
leso <- read_rds("data-processed/leso.rds")
austin_msa_agencies <- read_csv("data-raw/austin_msa_agencies.csv")

cols <- c(
  "agency_name",
  "item_name",
  "quantity",
  "ui",
  "acquisition_value",
  "ship_date"
  )
```

## Glimpse the data


```{r}
leso %>% glimpse()
```


## Filter to TX

```{r tx}
tx <- leso %>% 
  filter(state == "TX")
```

## Central Texas

This is a list of agencies from utdata/rwd-police-agencies for comparison.


```{r show_agencies}
austin_msa_agencies %>% 
  arrange(agency)
```

### Make a list of Austin MSA agencies

This is a very suspect list because it comes from looking through the available agencies "by hand" from the Sept 2020 data. If an item has `#NC` by it that means it has not been confirmed by a record in the LESO data and may not match. Does not look for public school districts.

Last checked for new items in 04012024 release.

```{r austinmsa}

aus_msa_list <- c(
  "AUSTIN PARKS POLICE DEPT", #NC
  "AUSTIN POLICE DEPT",
  "BASTROP COUNTY SHERIFF'S OFFICE",
  "BASTROP POLICE DEPT",
  "BEE CAVE POLICE DEPT",
  "BUDA POLICE DEPT",
  "CALDWELL COUNTY SHERIFFS OFFICE",
  "CEDAR PARK POLICE DEPT",
  "ELGIN POLICE DEPARTMENT",
  "FLORENCE POLICE DEPT", #NC
  "GEORGETOWN POLICE DEPT",
  "GRANGER POLICE DEPT", #NC
  "HAYS CO CONSTABLE PRECINCT 4",
  "HAYS COUNTY SHERIFFS OFFICE",
  "HUTTO POLICE DEPT",
  "JARRELL POLICE DEPT", #NC
  "JONESTOWN POLICE DEPT", #NC
  "KYLE POLICE DEPT",
  "LAGO VISTA POLICE DEPT",
  "LAKEWAY POLICE DEPT", 
  "LEANDER POLICE DEPT",
  "LIBERTY HILL POLICE DEPT", #NC
  "LOCKHART POLICE DEPT",
  "LULING POLICE DEPT",
  "MANOR POLICE DEPT",
  "MARTINDALE POLICE DEPT", #NC
  "PFLUGERVILLE POLICE DEPT",
  "ROLLINGWOOD POLICE DEPT", #NC
  "SAN MARCOS POLICE DEPT",
  "SMITHVILLE POLICE DEPT", #NC
  "SUNSET VALLEY POLICE DEPT", #NC
  "TAYLOR POLICE DEPT", #NC
  "THRALL POLICE DEPT", #NC
  # TEXAS STATE UNIVERSITY HI_ED
  "TRAVIS COUNTY SHERIFFS OFFICE",
  # TRAVIS CONSTABLE OFFICE,
  # SOUTHWESTERN UNIVERSITY HI_ID
  "WESTLAKE HILLS POLICE DEPT", #NC
  "UNIV OF TEXAS SYSTEM POLICE HI_ED",
  "WILLIAMSON COUNTY SHERIFF'S OFFICE"
)

aus_msa <- tx %>% 
  filter(agency_name %in% aus_msa_list)
```


For running checks

```{r stations}
tx %>% 
  count(agency_name)
```


### Austin MSA

```{r}
aus_msa %>% 
  group_by(agency_name) %>% 
  summarize(
    cnt = n(),
    total_value = sum(acquisition_value * quantity)
  ) %>% 
  arrange(total_value %>% desc())
```

### Williamson County Sheriff

```{r wilco}
wilco <- aus_msa %>% 
  filter(agency_name == "WILLIAMSON COUNTY SHERIFF'S OFFICE")

wilco %>% 
  select(cols) %>% 
  group_by(item_name) %>% 
  summarize(
    cnt = sum(quantity),
    total_value = sum(acquisition_value * quantity)
    ) %>% 
  arrange(total_value %>% desc())
```

```{r}
wilco %>% 
  filter(
    item_name == "RIFLE,7.62 MILLIMETER"
  ) %>% 
  select(cols)
```

## Texas higher education

Look for K-12 and HI_ED

```{r}
tx %>% 
  filter(
    str_detect(agency_name, "HI_ED")
  ) %>% 
  group_by(agency_name) %>% 
  summarise(
    cnt = sum(quantity),
    total_value = sum(quantity * acquisition_value)
  ) %>% 
  arrange(total_value %>% desc())
```

## Texas local school districts

```{r}
tx %>% 
  filter(
    str_detect(agency_name, "K-12")
  ) %>% 
  group_by(agency_name) %>% 
  summarise(
    cnt = sum(quantity),
    total_value = sum(quantity * acquisition_value)
  ) %>% 
  arrange(total_value %>% desc())

```

### Spring Branch

```{r}
tx %>% 
  filter(
    agency_name == "SPRING BRANCH ISD PD K-12"
  ) %>% 
  group_by(item_name) %>% 
  summarise(
    cnt = sum(quantity),
    total_value = sum(quantity * acquisition_value)
  )
```

Checking pistol

```{r}
tx %>% 
  filter(
    agency_name == "SPRING BRANCH ISD PD K-12",
    item_name == "PISTOL,CALIBER .45,AUTOMATIC"
  ) %>% select(cols, -agency_name)
```

